\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2024-04-19 paper.tex

% Para 2: What are they, though?
Although it is widely accepted that ARGs are important, there is some
confusion about what, precisely, an ARG \emph{is}.
% The grammar is a nightmare here - the stochastic process is singular, right?
In its original form,
developed by Griffiths and colleagues,
the ARG is an alternative
formulation of the coalescent with recombination~\citep{hudson1983properties},
where the stochastic process of coalescence and recombination
among ancestral lineages is formalised
as a graph~\citep{griffiths1991two,
ethier1990two,griffiths1996ancestral,griffiths1997ancestral}.
% Move into plural, an ARG is an instance of the data structure in this view
Subsequently, an ARG has come to be thought of as a data
structure~\citep{minichiello2006mapping}, i.e.\ describing
a \emph{realisation} of such a random process,
or an inferred ancestry of a sample of genomes.
The distinction between stochastic process
and data structure is not clear cut, however, and subfields use the term
differently (Appendix~\ref{sec-arg-history},\ref{sec-big-and-little-arg}).
% % TODO rework these two sentences. Point about models vs structures 
% % is an important one, we need to make that.
% Recent large-scale methods infer approximate \emph{structures}
% instead of a complete and fully detailed history
% (Appendix~\ref{sec-survey-arg-infer}), and it is
% therefore important to distinguish between
% approximate structures and approximate models.
% Whether an inference method is based
% on heuristics or a rigorous mathematical model is orthogonal
% to the level of detail provided in its estimate.
The term ``ARG'' therefore has many different meanings, varying 
over time and depending on context. 
There is, however, an emerging consensus to use the term
in quite a general sense~\citep[e.g.][]{mathieson2020ancestry,hejase2020summary,
schaefer2021ancestral,harris2023using,zhang2023biobank,fan2023likelihood},
informally
encompassing the varied structures output by modern simulation and
inference methods~\citep{rasmussen2014genome, palamara2016argon, haller2018tree,
kelleher2019inferring, speidel2019method, baumdicker2021efficient,
zhang2023biobank}. 
However, there is currently no formal definition or systematic 
discussion that unifies these different structures, 
ultimately stifling progress in this vibrant research area.

>>>>

% Para 2: What are they, though?
Although it is widely accepted that ARGs are important, there is some
confusion about what, precisely, an ARG \emph{is}.
% The grammar is a nightmare here - the stochastic process is singular, right?
In its original form,
developed by Griffiths and colleagues,
the ARG is an alternative
formulation of the coalescent with recombination~\citep{hudson1983properties},
where the stochastic process of coalescence and recombination
among ancestral lineages is formalised
as a graph~\citep{griffiths1991two,
ethier1990two,griffiths1996ancestral,griffiths1997ancestral}.
% Move into plural, an ARG is an instance of the data structure in this view
Subsequently, an ARG has come to be thought of as a data
structure~\citep{minichiello2006mapping}, i.e.\ describing
a \emph{realisation} of such a random process,
or an inferred ancestry of a sample of genomes.
The distinction between stochastic process
and data structure is not clear cut, however, and subfields use the term
differently (Appendix~\ref{sec-arg-history},\ref{sec-big-and-little-arg}).
% % TODO rework these two sentences. Point about models vs structures 
% % is an important one, we need to make that.
% Recent large-scale methods infer approximate \emph{structures}
% instead of a complete and fully detailed history
% (Appendix~\ref{sec-survey-arg-infer}), and it is
% therefore important to distinguish between
% approximate structures and approximate models.
% Whether an inference method is based
% on heuristics or a rigorous mathematical model is orthogonal
% to the level of detail provided in its estimate.
The term ``ARG'' therefore has many different meanings, varying 
over time and depending on context. 
There is, however, an emerging consensus to use the term
in quite a general sense~\citep[e.g.][]{mathieson2020ancestry,hejase2020summary,
schaefer2021ancestral,harris2023using,zhang2023biobank,fan2023likelihood},
informally
encompassing the varied structures output by modern simulation and
inference methods~\citep{rasmussen2014genome, palamara2016argon, haller2018tree,
kelleher2019inferring, speidel2019method, baumdicker2021efficient,
zhang2023biobank}. 
There is currently no formal definition or systematic 
discussion that unifies these different structures,
however, stifling progress in this vibrant research area.
